Species persistence under climate and fishing

### 2014-01-17 Writing/MS.tex
working on the discussion

In our model, the adults from the current year produce offspring according to a recruitment function, and these 
offspring disperse across the one-dimensional world according to a dispersal kernel to become the next 
generation's adults. The adults are subjected to harvesting before they produce offspring so that only a 
proportion of the fish survive to reproduce. We incorporated these processes-- recruitment, harvesting, and 
dispersal-- into an integrodifference model to describe how the population changes over time. If $n_t(x)$ 
is the density of fish at position $x$ at time $t$, then the density of fish at the next generation is given by

@@

We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist. 
We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
We can find these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
source of larvae and its dependence on the destination of the larvae,
i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.  In our analyses, we used one such separable kernel, the Gaussian kernel given by

@@

\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.  To quantify the effects of both a shifting environment and harvesting pressure, we find the total biomass in the equilibrium traveling wave. As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

@@

We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
{Synergy}).  In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.  The stressors interact most strongly when they are both high, shortly before they drive the population extinct.  Our simulations with a Laplace kernel produce similar 
results. In the Appendix, we show analogous results for a sinusoidal dispersal kernel.

@@

We found that thresholds of any amount alleviate the effects of harvesting, and the ability of the population to 
persist is recovered (Figure \ref{thresholds}). Thus when thresholds are in place, the harvesting rate no longer 
determines the critical rate of environmental shift $c^*$. We also examined the effect of marine protected 
areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and 
climate change parameters where the fish population could survive. With MPAs in place, the population had a 
slightly higher abundance along the edges of the patch where the population is limited by harvesting, which 
translated into a slightly increased critical harvest rate (Figure \ref{MPAs}). Additionally MPAs increased 
overall catch at the highest harvest rates under which the population could survive (Figure \ref{MPAs}).

@@

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$. }
\label{thresholds}
\end{figure}

@@

We extended the model of Zhou and Kot \citep{ZhouKot2011} to study a fish population subject to both the climate-driven range shifts already included and harvesting. Previous work has highlighted the importance of 
reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]
{Hastingsetal2005} and the integrodifference model explicitly includes these processes. We found that 
the two disturbances interact synergistically with each other in some cases. However, we also found that 
placing thresholds on harvesting intensity can alleviate interactions between the two stressors and that spatial 
management increased the maximum harvesting rate at which the population could survive.

We found that the synergy between the two stressors is positive and that the synergy is greatest in the region 
of parameter space where the equilibrium biomass is smallest. Additionally, we found that the higher the 
growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a 
population can adjust to harvest and climate change. We find similar results from the analytically derived 
biomass and the simulation derived biomass. This indicates that this result is robust to changes in the 
dispersal kernel. Similarly, in experiments, \cite{Moraetal2007} found no interaction between harvesting and 
habitat fragmentation but did find an interaction between harvesting and warming temperature \citep
{Moraetal2007}. We chose to measure the effect of each stressor by the drop in biomass caused by the 
stressor, as in \cite{Crainetal2008}. In general, measuring synergy additively is more conservative than 
measuring synergy multiplicatively: stressors that have positive multiplicative synergy may have zero additive 
synergy \citep{Crainetal2008, Foltetal1999}. Since we found small levels of positive additive synergy between 
the two stressors, other measures of synergy might show even higher levels of interaction. This synergy and 
the fact that it most severely affects those populations whose persistence is most tenuous is worrisome from a 
conservation perspective. Harvesting levels or rate of environment shift that are sustainable individually 
together can drive a population to extinction. It is therefore important to take the effects of both stressors into 
consideration when designing conservation and management strategies.

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the 
population's biomass at equilibrium and improved its ability to persist. These management strategies also 
affect how the two stressors interact with each other.  Protected areas have 
been advanced as a way to help organisms keep pace with range shifts as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, 
Hannahetal2007,Botsfordetal2001, Gaylordetal2005, 
HastingsBotsford2003,Thomasetal2012}.  While MPAs only change harvesting practices 
and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting 
and range shifts will help to better implement and place protected areas.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do 
not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist 
over time. However, we found that qualitatively similar results about the interaction between climate and 
harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age 
structure in our model. The effects of both harvesting and climate change may be different across different age 
classes; including this level of complexity is left for future work. Additionally, we did not include any 
mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.

In addition to these species-specific extensions, this modeling framework can easily be adjusted to consider 
species interactions, especially predator-prey pairs. Predatory species affect marine populations and 
considering a set of species that included both predators and prey would be equivalent to imposing additional 
stressors on the prey species whose interactions we would like to measure \citep{Lingetal2009, 
Gurevitchetal2000}. Considering multiple species would not only allow us to understand the effects of multiple 
stressors on prey populations but would also afford us the opportunity to study how community composition 
would change as a result of climate change, as has been observed in some marine systems \citep
{Urbanetal2012}.

>>>>

In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function, and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.  We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce. These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time. If $n_t(x)$ 
is the density of fish at position $x$ at time $t$, then the density of fish at the next generation is given by

@@

We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist. 
We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
We found these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
source of larvae and its dependence on the destination of the larvae,
i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.  In our analyses, we used one such separable kernel, the Gaussian kernel given by

@@

\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.  To quantify the effects of both a shifting environment and harvesting pressure, we found the total biomass in the equilibrium traveling wave. As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

@@

We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
{Synergy}).  In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.  The stressors interact most strongly when they are both high, shortly before they drive the population extinct.  We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results.

@@

We found that thresholds of any amount alleviate the effects of harvesting, and the ability of the population to 
persist is recovered (Figure \ref{management}). Thus when thresholds are in place, the harvesting rate no longer 
determines the critical rate of environmental shift $c^*$. We also examined the effect of marine protected 
areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and 
climate change parameters where the fish population could survive. With MPAs in place, the population had a 
slightly higher abundance along the edges of the patch where the population is limited by harvesting, which 
translated into a slightly increased critical harvest rate (Figure \ref{management}). Additionally MPAs increased 
overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).

@@

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$. }
\label{management}
\end{figure}

@@

We extended a spatial to study a fish population subject to both the climate-driven range shifts and harvesting. Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.  In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.  For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.

%% discussion of critical rates and curved persistence boundary

To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.  We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.  We found similar results from the analytically derived biomass and the simulation derived biomass. This indicates that this result is robust to changes in the dispersal kernel.  We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.  We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.  In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.  Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.

While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e. synergistic effects were quite small.  However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.  Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.  Synergy between harvesting and climate changes, e.g. warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.  Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.  Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.  Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.  The management strategies also affect how the two stressors interact with each other.  Specifically, harvesting thresholds can alleviate interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work. Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
